### 2011-10-28 inst/doc/article.Rnw
Added article and manual Rnw files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The qcl mapping routine
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article}

\usepackage{times}
\usepackage{amsmath}
\usepackage{url}
\usepackage{color}

% revise margins
\setlength{\headheight}{0.0in}
\setlength{\topmargin}{-0.25in}
\setlength{\headsep}{0.0in}
\setlength{\textheight}{9.00in}
\setlength{\footskip}{0.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}

\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

\newcommand{\code}{\texttt}
\newcommand{\lod}{\text{LOD}}

% make "Figure" within figure captions a small font
\renewcommand{\figurename}{\small Figure}

\begin{document}

\SweaveOpts{prefix.string=Figs/fig}
\setkeys{Gin}{width=\textwidth} %% <- change width of figures
% Try to get the R code from running into the margin
<<echo=FALSE>>=
options(width=87, digits=3, scipen=4)
set.seed(61777369)
@

% Change S input/output font size
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0, 0, 0.56}}}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0.56, 0, 0}}}

\begin{center}
\vspace*{1in}
\textbf{\Large QCL mapping in experimental crosses}

\bigskip \bigskip 
\bigskip \bigskip

{\large Danny Arends, and Ritsert C. Jansen}

\bigskip \bigskip

{\color{blue} \tt \bfseries University of Groningen \\
Groningen Bioinformatics centre \& GCC \\
Technical Report \# 214 \\[24pt]
XX Okt 2011}

\end{center}

\bigskip \bigskip

\textbf{\sffamily Abstract}: Quanitative correlation locus (QCL) mapping is 
a novel approach to detect genetic regulation of phenotypes in recombinant 
inbred line populations (RIL)
\vfill

\noindent \emph{email}: \verb|Danny.Arends@Gmail.com|
\newpage

\textbf{\sffamily Introduction}

Some introduction text

\bigskip

\textbf{\sffamily Methodology}
\nopagebreak

Some methodology text

\bigskip

\textbf{\sffamily Results}
\nopagebreak

Some resulting text

\bigskip
\textbf{\sffamily Discussion}
\nopagebreak

Some discussion text

\newpage
{\small
\noindent \textbf{\sffamily Package qcl function overview} \\
\noindent \begin{tabular}{ll} 
\hspace*{25mm} & \hspace*{103mm} \\ \hline
QCLscan & Genome wide scan for QCL \\
QCLscan.network & Scan for QCL using start gene(s) \\
QCLpermute & Significance thresholds for QCL by permutation \\
plot.QCL & Plot QCL profiles at increasing cut-offs \\
image.QCL & Image a multiple phenotype scan \\
\hline
\end{tabular}
}

\end{document}