### 2011-11-10 inst/doc/article.Rnw
Added the science style + updates to the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The qcl mapping routine
%\VignetteIndexEntry{Article}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article}

\usepackage{scicite}
\usepackage{times}
\usepackage{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}

% revise margins
\setlength{\headheight}{0.0in}
\setlength{\topmargin}{0.0cm}
\setlength{\headsep}{0.0in}
\setlength{\textheight}{21.0cm}
\setlength{\footskip}{1.0cm}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{16.0cm}

\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

\newcommand{\code}{\texttt}
\newcommand{\lod}{\text{LOD}}

% make "Figure" within figure captions a small font
\renewcommand{\figurename}{\small Figure}

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

\title{QCL: Added value of differential correlation}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Danny Arends,$^{1\ast,2}$ Ritsert C. Jansen,$^{1}$\\
\\
\normalsize{$^{1}$Groningen Bioinformatics Centre - University of Groningen}\\
\normalsize{Nijenborgh 7, 9747 AG Groningen, The Netherlands}\\
\normalsize{$^{2}$Genomics Coordination Centre - University Medical Centre Groningen}\\
\normalsize{Hanzeplein 1, 9700 RB Groningen, The Netherlands}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  Danny.Arends@Gmail.com.}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

\SweaveOpts{prefix.string=fig}
\setkeys{Gin}{width=\textwidth} %% <- change width of figures
% Try to get the R code from running into the margin
<<echo=FALSE>>=
options(width=87, digits=3, scipen=4)
@

<<echo=FALSE,results=hide>>=
  require(qcl)
  set.seed(61777369)
  data(ath.metabolites) # Arabidopsis Thaliana dataset
  qcl_result <- QCLscan(ath.metab$genotypes, ath.metab$phenotypes)
@

% Change S input/output font size
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0, 0, 0.56}}}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0.56, 0, 0}}}

\begin{sciabstract}
In this paper we present Quanitative Correlation Locus (QCL) mapping, a novel approach 
to detect genetic regulation of phenotypes in recombinant inbred line populations (RIL) 
or other experimental crosses. QCL mapping allows association of phenotype correlation 
differences to genetic markers, in almost the same way as QTL mapping associates genetic 
markers with differences in expression.
We use QCL mapping to reconstruct underlying genetic architecture, and provide insight 
into regulation of (gene) expression. QCL information in combination with classical QTL 
information leads to a more detailed insight into the genetic architecture underlying 
complex traits. The method also provides a vital link in regulatory network reconstruction.
\end{sciabstract}

\section*{Introduction}

Genetical genomics experiments have shown that gene expression variation can be mapped to 
genotypic variation \cite{Jansen:2001}. Quantitative Trait Locus (QTL) mapping of a 
gene identifies regions in the genome for which different genotypes lead to differential 
expression level of the gene. This QTL mapping can be done on all molecular levels including, 
gene expression (eQTL), protein abundance (pQTL) and metabolites (mQTL).

Recently different methods have been proposed that allow studying, in addition to the genes 
differentially expressed between conditions, the genes that are differentially coexpressed 
between conditions \cite{Tesson:2010,Kompass:2011}. Such differential co-expression 
is taken as another level of information from genomic data. This level of information can for 
example be used to expose transcriptional network rewiring \cite{Langfelder:2008}. Differential 
co-expression has been observed in studies of different diseases, including between healthy and 
cancer tissue studies. In parallel, correlations between phenotypes are studied for various 
reasons e.g. discovering and detecting batch effects, marker assisted breeding and causal 
inference. The importance of identifying differences in correlation has been increasingly 
recognized and this led to the development of this algorithm to perform QCL analysis in 
experimental crosses.

Using recombinant inbred lines we introduce natural variation into a population 
detectable as cis-QTL in a genome wide QTL scan. The effect of any such genetic 
variation is transfered downstream to interaction partners. When this leads to 
changes in mean expression of one of the partners we can also detect this using 
classical QTL methods (trans/distant QTL).

However QTL mapping suffers from two major drawbacks (p1) the interaction partners 
of genes are obscured by many phenotypes co-localizing at a single genetic marker, and 
(p2) due to buffering the effect of any variation will 'fade' quickly and not be 
detectable as the difference in mean phenotype in the population does not change 
significantly.

We proposed to map not differential co-expression but differential correlation to genetic 
variation, i.e. to identify regions in the genome for which one genotype value leads to 
correlated expression between a start gene and all other genes, exploiting the fact that any 
cis-QTL will induce variation which will in turn lead to correlation differences among 
individuals carrying the other genotype. We merge information coming from QCL and QTL to reveal 
underlying genetic architecture, which would remain hidden when using only classical QTL information.

We urge all QTL researchers to combine QCL analysis with existing QTL analysis to expose 
transcriptional networks involved in the regulation of complex phenotypes. For this we 
provide an R package with a reference implementation  of the QCL algorithm

\bigskip

\section*{Methodology}
\nopagebreak

\emph{\sffamily Preprocessing}

Probes showing little to no variance (< 0.01) were removed from the analysis. Batch 
effects due to different hybridization dates were estimated and removed from the data 
by using a mean centering approach (Group means were estimated and corrected).

\section*{Algorithm: QCL mapping}

QCL mapping however uses this buffering to detect interaction partners. The hypothesis 
is that when we split the population based on genotype at the cis-marker, related phenotypes 
will loosing correlation with the gene under investigation. This allows us to not only 
only identify its interaction partners. QCL mapping is thus not suffering from (p1) many 
phenotypes co-localizing). Additionally it is more sensitive to detect variation caused by 
regulators genes because variation of these genes causes many 'downstream' phenotypes 
to show differential correlation with the inducing phenotype, because QCL relies on 
correlation changes as opposed to differences in phenotype mean QCL is unaffected by (p2).

\emph{\sffamily Choosing a start/seed probe (Optional)}

Choose a phenotype to subject to QCL analysis, this can be your favourite gene, 
protein, metabolite or classical phenotype.

\emph{\sffamily Single phenotype QCL}

Using genotype data we can split individuals at each marker in two groups (A or B).
At each marker calculate the correlation between our phenotype and all other 
phenotypes in groups A and B. Collect all the result vectors into two matrices A 
and B, and calculate the QCL matrix by taking the squared difference between matrix A 
and B.

\emph{\sffamily Significance testing}

The threshold for a significant difference in QCL is assessed by using permutations, in 
each round the link between genotype and phenotype is broken, by redistributing at random 
genotypes amongst individuals not allowing for duplicates. The QCL on these new 
individuals allows for estimation of the underlying null-distribution of QCL scores. 
See figures XX, and XX showing the score distribution in permuted data.

\emph{\sffamily Creating single trait QCL profiles}

After permutation single trait QCL profiles can be obtained by setting a qcl significance 
threshold. This threshold defines when a change in correlation is significant, and at 
each marker we can plot the count of number the other phenotypes that show a significant 
difference in correlation. Overplotting chromosome, QTL information, and phenotype location 
(if applicable) produces the single QCL plots seen in figures XX, and XX.

\emph{\sffamily Step 3 Network scan for QCL}

If we find any significant QCL information we can try to reconstruct the interaction 
network of the phenotype. We first do a single phenotype QCL scan (step 2), then 
using a permutation determin a qcl threshold, and select phenotypes which show a 
significant change in QCL.

Two things can occur: 
1) No genes show a significant change in QCL, which means we 
cannot find any interaction partners for this phenotype, or 
2) This genes has an QCL with other genes, then the algorithm adds 
these genes onto a list to scan and repeats the single phenotype 
QCL scan (step 2) for these phenotypes (not allowing for duplicate 
scans). This is repeated until no more significant QCL are found or 
a user defined search depth is reached.

\emph{\sffamily Visualizing the Phenotypes x Marker relations}

When plotting multiple QCL profiles as a heatmap, multiple QCL are seen mapping to the 
same genetic location. At these locations, we expect to find one (or more) of the genes 
to show a cis-QTL, this is because a QTL can induce a QCL, because variation of expression 
in the phenotype is causing other phenotypes to become uncorrelated, when assigned to 
groups determined by the QTL marker. This phenotype (or phenotypes) introduced genetic 
variation, which is seen also (and more pronounced) in a QTL scan.

\emph{\sffamily Visualizing the Phenotypes x Phenotypes relations}

The additional information obtained by QCL mapping is summarized in the 
Phenotypes x Phenotypes matrix. This shows the summed evidence of interaction 
between the phenotypes showing QCL transmitted through all the markers and the phenotype scanned.

\emph{\sffamily Network reconstruction}

So how do we build up gene interaction networks, This looks complicated but the information 
is implicitely stored in the Phenotype x Marker and Phenotype x Phentype relations matrices. 
By using network scanning we only identify part of the network in which our phenotype is 
active and closely related interaction partners. To identify a pathway or main regulatory phenotypes 
controlling our phenotype we need to combine this information.

We show two examples in which we reconstruct the genetic architecture using QCL mapping. 
Both datasets are available inside the R package, as well as the code to analyse these 
two datasets.

\subsection*{Materials}
\emph{\sffamily Dataset Example 1. A. Thaliana}

Metabolite data from an A. Thaliana Landsberg erecta (Ler) and Cape Verde Islands (Cvi) 
dataset from R/qtl, 24 characterized metabolites which are the products of chain elongation. 
There are 162	recombinant inbred lines genotyped at 117 markers on 5 chromosomes. Additional 
details about the experimental settings are available in the original paper \cite{Keurentjes:2006,AlonsoBlanco:1998}

\emph{\sffamily Dataset Example 2. S. Cerevisiae}

S. Cerevisiae dataset from the gene expression omnibus (GEO) \cite{Edgar:2002} (GEO 
accession GDS1115, and GDS1116, \cite{Brem:2005}. This dataset contains expression 
profiles of parental strains and haploid progenies from a cross of strain BY4716 and 
the wild wine strain RM11-1a. Data was collected for 4482 probes. Genotype data 
for the 109 progenies was available at 282 genetic markers obtained by contacting 
the authors of the original paper. 
Expression of probes was measured under growth conditions containing 1\% ethanol. 
Additional details about the experimental settings are available in the original 
paper \cite{Brem:2005}.

\bigskip
\section*{Results}
\nopagebreak

\emph{\sffamily Example 1. A. Thaliana}

QCL mapping was performed on 24 characterized metabolities measured in Arabidopsis 
Thaliana. A significance threshold for QCL mapping was determined by using 1000 
permutations, the outcomes can be seen in table XX. We opted for the stringent 
threshold at P < 0.001 to use as 'significant' QCL score for further analysis.

Overview plots of the results of the QCL mapping in Arabidopsis can be found in 
Figure XX Phenotype to Marker interaction, which is comparable to the QTL mapping 
(Fig. XX). The additional information obtained by QCL mapping is shown in a heatmap 
of the Phenotype to Phenotype interaction matrix (figure XX) and the same information 
but now represented as an interaction network is seen in figure XX
\begin{table}[ht]
\begin{tabular}{ l | c c }
P-value & QCL threshold Arabidopsis & QCL threshold Saccharomyces\\
\hline
5 \% & 0.220 & 0.435\\
1 \% & 0.294 & 0.495\\
0.1 \% & 0.375 & 0.562\\
\end{tabular}
\caption{QCL thresholds obtained after 1000 permutations}
\label{permutations}
\end{table}

\emph{\sffamily Example 1. S. Cerevisiae}

QCL mapping was performed on

\bigskip
\section*{Discussion}
\nopagebreak

Microarray experiments have allowed the identification of genes differentially 
expressed between various conditions, or groups of genes coexpressed across 
samples.

Many genetical genomics screens have been carried out to map differential 
expression to genetic variation. Screening the genome for genotype combinations 
for which there is differential expression has been a successful strategy when 
mapping Mendelian traits. However results on complex traits have been less 
then satisfactory with only small amounts of heritability explained by 
individuals loci. 
QCL mapping however 'borrows' information from each phenotype in the analysis set, 
and enables extraction of phenotype to marker to phenotype networks. 
Furthermore QCLs seem to 'transfer' from phenotype to phenotypes in a stabile 
fashion, which gives greater detail and confidence when reconstructing genetic 
architecture. 
It can provides better insight in the missing of heritability of variation, we 
hypothesize that changes in variation are affecting (by loss of) correlation in 
the normally correlated phenotypes leading to an amplification of this variation, 
by distrupting the network environment of which endo-phenotypes are part of.

\section*{Acknowledgements}
This work was supported by the Centre for BioSystems Genomics (CBSG) and the 
Netherlands Consortium of Systems Biology (NCSB), both of which are part of the Netherlands 
Genomics Initiative / Netherlands Organisation for Scientific Research [to DA], and 
by the EU 7th Framework Programme under the Research Project PANACEA, [222936 to RJ].

Also we'd like to mention Joeri van der Velde and Konrad Zych for supporting me with some very 
valuable feedback on figures presented here.

We'd like to thank readers for their attention, and invite them to 
address any questions to Danny Arends, at Danny.Arends@gmail.com.

\newpage
\bibliography{scibib}
\bibliographystyle{Science}

\newpage

\section*{Figures}

\begin{figure}[ht]
  \centerline{\includegraphics{QCL.pdf}}
  \caption{QCL mapping: The underlying hypothesis of QCL mapping. Seen here is a network of 3 phenotypes, in 
  which expression of P1, P2 and P3 is measured, and we want to infer the hidden relation between P1 
  and P2. In our model a genetic pertubation is causing P1 to show a QTL at a certain marker. We expect variation 
  induced by P1 to transfer to P2, However (like in most biological systems) expression of P2 is tightly controlled, 
  and variation seen in phenotype 1 wil not induce a QTL in phenotype 2. However when we partition individuals based 
  on the genotype at the QTL marker of P1, we can make use of the correlation structure between P1 and P2. Allowing 
  us to infer the relationship between phenotype 1 and 2, when we see difference in correlation between P1A and P2A 
  and P1B and P2B. We don't observe this difference in correlation between P3 and P1, meaning there is no relation 
  between P1 and P3}
  \label{QCL_img}
\end{figure}

\begin{figure}
  \centering
<<singletrait, fig=TRUE, echo=FALSE, height=3.5>>=
plot(qcl_result, 2, do.legend=FALSE)
@ 
  \caption{Single trait QCL profile (A. Thaliana) shows the number of phenotypes that have a significant 
  QCL with the scanned phenotype at different QCL significance thresholds.\label{fig:singletrait}}
\end{figure}

\begin{figure}[ht]
  \centering
<<p2mmatrix, fig=TRUE, echo=FALSE, height=3.5>>=
par(mar=c(3,8,1.1,1.1))
image(qcl_result, 0.5, against="markers",do.grid=FALSE)
@ 
  \caption{Phenotypes x Markers matrix (A. Thaliana) this is 'comparable' to a QTL scan on multiple traits.\label{fig:p2mmatrix}}
\end{figure}

\begin{figure}[ht]
  \centering
<<p2pmatrix, fig=TRUE, echo=FALSE, height=3.5>>=
par(mar=c(3,8,1.1,1.1))
image(qcl_result, 0.5, against="phenotypes")
@ 
  \caption{Phenotypes x Phenotypes matrix (A. Thaliana) additional information obtained by QCL mapping, 
  which can be used to reconstruct the underlying genetic architecture.\label{fig:p2pmatrix}}
\end{figure}

\begin{figure}[ht]
  \centerline{\includegraphics{AraNet.pdf}}
  \caption{Reconstructed network for chain elongation in Arabidopsis Thaliana}
  \label{AraNet}
\end{figure}

\begin{figure}[ht]
  \centerline{\includegraphics{AraPerm.pdf}}
  \caption{Histogram of scores obtained durting permutation of the QCL mapping methodology}
  \label{AraPerm}
\end{figure}

\end{document}