### 2012-01-30 inst/methodPaper/document.tex
Phenotype -> trait

\documentclass{bioinfo}
\usepackage{subfigure}
\usepackage{amsmath}

\copyrightyear{2012}
\pubyear{2012}

\begin{document}
\firstpage{1}
\title[CTLmapping]{CTL mapping - The added value from segregating correlation}
\author[Arends \emph{et al.}]{
  Danny Arends\,$^{1,\footnote{to whom correspondence should be addressed}}$,
  and Ritsert C. Jansen\,$^{1}$
}
\address{
  $^{1}$Groningen Bioinformatics Centre, University of Groningen, Groningen, The Netherlands.
}
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
\editor{Associate Editor: XXXXXXX}
\maketitle
\begin{abstract}
  \section{Motivation:}
  Here we present the concept of Correlated Traits Locus (CTL) mapping, a novel approach 
  to detect genetic regulation of traits in recombinant inbred line populations (RIL) 
  or SNP genotypes populations. CTL mapping allows association of trait correlation 
  differences to genetic markers, in almost the same way as QTL mapping associates genetic 
  markers with differences in expression.

  We use CTL mapping to reconstruct underlying genetic architecture, and provide insight 
  into regulation of (gene) expression. CTL information in combination with classical QTL 
  information leads to a more detailed insight into the genetic architecture underlying 
  complex traits. CTL mapping provides a missing link in regulatory network reconstruction.

  \section{Results:}
  The software is available under the terms of GNU General Public License, version 3
  (\href{http://www.r-project.org/Licenses/GPL-3}{www.r-project.org/Licenses/GPL-3}).
  \section{Availability:}
  Software, documentation and source code are available from:
  \href{http://www.mapctl.org}{www.mapctl.org}
  \section{Contact:} 
  \href{Danny.Arends@Gmail.com}{Danny.Arends@Gmail.com}
\end{abstract}

\section{Introduction}
  Genetical genomics experiments have shown that gene expression variation can be mapped to 
  genotypic variation \cite{Jansen:2001}. Quantitative Trait Locus (QTL) mapping of a 
  gene identifies regions in the genome for which different genotypes lead to differential 
  expression level of the gene. This QTL mapping can be done on all molecular levels including, 
  gene expression (eQTL), protein abundance (pQTL) and metabolites (mQTL).

  We present Correlated Traits Locus (CTL) mapping as a comparable method to QCL mapping, 
  CTL unlike QTL associates differential correlation to genetic variation, i.e. to identify regions in the 
  genome for which one genotype value leads to correlated expression between a trait and all 
  other traits, exploiting the fact that any QTL will be the result of either direct induced 
  variation or indirect variation, which will be visible as correlation differences among 
  individuals carrying different genotypes.

  We merge information coming from CTL and QTL to reveal underlying genetic architecture, which 
  would remain hidden when using only classical QTL information.

\section{Methodology}
  Choose a trait to subject to CTL analysis, this can be your favourite gene, 
  protein, metabolite or classical trait.

  Split the population at a genetic marker in two groups (aa and bb). Calculate the correlation 
  between our trait of interest and other traits. Calculate the CTL matrix by taking the 
  squared difference between matrix A and B.

  Single trait CTL profiles can be obtained by transforming the CTL scores into LOD scores. 
  For this we use permutations obtained during 15.000 permutation. We note that for each trait this 
  shows a different distribution of permutation scores. We assess the likelyhood (per trait) and 
  take the $-log10(P)$ of every CTL score to compare QTL profiles with summarized CTL profiles. 
  A CTL profile show the contribution of each trait to the total likelyhood.

  \begin{figure*}[ht]
    \begin{center}
    \vspace{-0.2cm}
    \hspace{-1.3cm}
    \subfigure[CTL profiles]{
      \includegraphics[keepaspectratio,scale=.26]{img/ctl_qtl.pdf}
      \label{subfiga}
    }
    \hspace{0.3cm}
    \subfigure[Traits x Traits interaction matrix]{
      \includegraphics[keepaspectratio,scale=.65]{img/ctl_pxp.pdf}
      \label{subfigb}
    }
    \hspace{0.3cm}
    \subfigure[Reconstructed network]{
      \includegraphics[keepaspectratio,scale=.17]{img/ctl_net.pdf}
      \label{subfigc}
    }
  \end{center}
  \begin{minipage}{7in}
  \label{newplots}
  \vspace{-0.7cm}
  \caption[Plots]{
  {\emph {\bf a}}: Example CTL + QTL profiles (red). Seen are the QTL profile on top of the CTL profile. 
  The first peak is a CTL caused by X4.Methylsulfinylbutyl and X5.Methylthiopenty showing a significant 
  difference in correlation, we don't observe a QTL for this location. 
  {\emph {\bf b}}: Traits x Traits interaction matrix (TxT matrix). Information obtained using CTL mapping, 
  showing for each trait the significance of the difference in correlation to all other traits. This 
  matrix can be transformed into .sif file to generate a network image. 
  {\emph {\bf c}}:  Reconstructed network, this shows the same information as the TxT matrix only in a more accessible way, we visualized our 
  network using cytoscape, the linewidth is the summarized interaction evidence.
 }
\end{minipage}
\vspace{-0.5cm}
\end{figure*}

\emph{ {\bf Network reconstruction}}
  To build up gene interaction networks, we take the Traits x Marker and Traits x Phentype 
  relations matrices.  From the TxM we can identify partners of our trait, and from the PxP we 
  identify the location of out trait in the pathway or main regulatory traits.

\emph{ {\bf Significance testing}}
  Comparing CTL profiles to QTL profiles CTL scores need to be converted to LOD scores. Significance 
  of CTL is assessed by using permutations, in each round the link between genotype and 
  trait is broken, by redistributing at random genotypes amongst individuals not 
  allowing for duplicates. The CTL on these new individuals allows for estimation of 
  the underlying 0-distribution of CTL scores. When observe CTL scores in 
  real data which are higher than any CTL score obtained during permutation we 
  use a generalized Pareto distribution (GPD) to estimate the extreme tail 
  of the distribution \cite{Knijnenburg:2009}, to get likelyhood estimates for the 
  extreme scores observed.

\emph{ {\bf Materials}}
  Metabolite data from an \emph{A. Thaliana} Landsberg erecta (Ler) and Cape Verde Islands (Cvi) 
  dataset from R/qtl, 24 characterized metabolites which are the products of chain elongation. 
  There are 162 recombinant inbred lines genotyped at 117 markers on 5 chromosomes. Additional 
  details about the experimental settings are available in the original paper \cite{Keurentjes:2006,AlonsoBlanco:1998}
\vspace{-0.6cm}
\section{Results}
  CTL mapping was performed and a per trait CTL thresholds of $P < 0.001$ was determined by using 15.000 permutations.
  Information obtained by CTL mapping is shown in the trait to trait interaction matrix (Fig. \ref{subfigb}).
  The CTL profiles were combined with QTL information, leading to the following observations \ref{subfiga}:\\
  1) Additional CTL found, where no QTL was present. Loss of variation in the genes 
  underneath the CTL are not causing variation differences in the observed trait, CTL is able to 
  detect the differences in correlation.\\
  2) CTL co-localized with QTL, in this situation CTL gives a detailed view of which traits 
  are contributing variation to the observed QTL.\\
  3) QTL with no CTL, we interpret this as the source of the variation. Analysis in this way of 
  all 24 Single trait CTL profiles combined with their QTL profiles allows us to reconstruct the 
  source of natural variation and follow it down the network.

\emph{ {\bf Network reconstruction}}
  Using the previously decribed method to draw the arrows between traits, we reconstruct the interaction 
  network amongst the 24 metabolities in Arabidopsis (Fig. \ref{subfigc}). We observe that our network is 
  composed of 2 parts which seem connected only by a just significant CTL from X3.Benzoyloxypropyl to 
  Quercetin.deoxyhexosyl.dihexoside.

\section{Conclusion \& Discussion}
  Microarray experiments have identified genes differentially expressed between various 
  conditions, or groups of genes coexpressed across samples. Genetical genomics screens 
  are carried out to map differential expression to genetic variation. Screening 
  the genome for genotype combinations for which there is differential expression is 
  a successful strategy when mapping Mendelian traits.

  Results on complex traits however have been less then satisfactory with only small 
  amounts of heritability explained by individuals loci. CTL mapping 'borrows'
  information from all traits in the analysis, and extracts trait to trait 
   information.

  Furthermore CTLs seem to 'transfer' from trait to traits in a stabile 
  fashion, allowing reconstructing of genetic architecture. CTL combined with QTL 
  data identify gene transcript introducing variation (by absense), while co-localizing 
  with QTL affected by upstream variation.

\paragraph{Funding\textcolon}
The Centre for BioSystems Genomics (CBSG) and the Netherlands Consortium of Systems 
Biology (NCSB), both of which are part of the Netherlands Genomics Initiative / 
Netherlands Organisation for Scientiﬁc Research [to DA]; the EU 7th Framework 
Programme under the Research Project PANACEA [222936 to R.J.].
\vspace{-0.7cm}
\bibliographystyle{natbib}
\bibliography{document}
\end{document}