### 2012-05-10 Rctl/inst/doc/manual.Rnw
[Rctl] Manual updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The ctl mapping routine
%\VignetteIndexEntry{Manual}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article}

\usepackage{times}
\usepackage{amsmath}
\usepackage{url}
\usepackage{color}

% revise margins
\setlength{\headheight}{0.0in}
\setlength{\topmargin}{-0.25in}
\setlength{\headsep}{0.0in}
\setlength{\textheight}{9.00in}
\setlength{\footskip}{0.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}

\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

\newcommand{\code}{\texttt}
\newcommand{\lod}{\text{LOD}}

% make "Figure" within figure captions a small font
\renewcommand{\figurename}{\small Figure}

\begin{document}

\SweaveOpts{prefix.string=fig}
\setkeys{Gin}{width=\textwidth} %% <- change width of figures
% Try to get the R code from running into the margin
<<echo=FALSE>>=
options(width=87, digits=3, scipen=4)
set.seed(61777369)
@

% Change S input/output font size
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0, 0, 0.56}}}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0.56, 0, 0}}}

\begin{center}
\vspace*{1in}
\textbf{\Large CTL mapping in R }

\bigskip \bigskip 
\bigskip \bigskip

{\large Danny Arends, and Ritsert C. Jansen}

\bigskip \bigskip

{\color{blue} \tt \bfseries University of Groningen \\
Groningen Bioinformatics centre \& GCC \\
revision \# 1 \\[24pt]
XX Okt 2011}

\end{center}

\bigskip \bigskip

\textbf{\sffamily Abstract}: Correlated Trait Loci (CTL) mapping using the R package 'ctl'
\vfill

\noindent \emph{email}: \verb|Danny.Arends@Gmail.com|
\newpage

\textbf{\sffamily Some words in advance}
\nopagebreak

Welcome reader to this tutorial, you'll be thrown into the world of network reconstruction 
using differential correlations. This methodology CTL mapping is not the easiest but the 
results can be of great help unraveling trait to trait interactions in all types of 
eukaryotes. In this tutorial I'll try to use as few difficult concepts as possible. But
some always $sneak$ in.\\
\\
Enjoy this tutorial, and good luck hunting that heritability.\\
Danny Arends\\
\\
PS. Use a good text editor\\
PPS. Use a repository / backup system

\bigskip

\textbf{\sffamily Background \& Methodology}
\nopagebreak

Some introduction on the methodology, skip this chapter first (DO come back later) if you 
have a dataset and just want to use the software.

\bigskip

\textbf{\sffamily Format and load your data}
\nopagebreak

\emph{Loading your Excel / CSV data}\\
Use the example excel file and simply replace the example data by your own then: 
  $$File->Save As->TODO$$

After you have stored the data as plain text open the files in a text editor (NOT WORD, for 
windows Notepad++ is a good option) and check to see if your data is stored correctly, if 
not reformat your data in both the genotype.csv and the phenotype.csv files to match the 
format below. {\color{red}\emph{\bf{NOTE: Individuals must match between the genotype and 
the phenotype file...}}} The phenotypes.csv file containing the genotype matrix is stored 
individuals x genetic marker:

\begin{tabular}{lrrrrr}
      & Marker1 & Marker2 & Marker3 & ... & MarkerN \\
Ind1  & 1 & 1 & 2 & ... & 2 \\
Ind2  & 1 & 1 & 2 & ... & 2 \\
Ind3  & 1 & 1 & 1 & ... & 2 \\
...   & ... & ... & ... & ... & ... \\
IndN  & 1 & 1 & 1 & ... & 2 \\
\end{tabular}

The phenotypes.csv file containing individuals x trait measurements:

\begin{tabular}{lrrrrr}
      & Trait1 & Trait2 & Trait3 & ... & TraitN \\
Ind1  & 5.8    & 11.0   & 9.0    & ... & 1.6 \\
Ind2  & 6.3    & 12.2   & NA     & ... & 1.3 \\
Ind3  & 5.1    & 11.1   & 12.3   & ... & 2.0 \\
...   & ...    & ...    & ...    & ... & ... \\
IndN  & 9.8    & 15.8   & 23.0   & ... & 3.4 \\
\end{tabular}

After verifying your files are formatted correctly (this is the source of most errors).
Start R and load in the data, check that the output looks like below:
<<echo=TRUE,eval=FALSE>>=
setwd("<PATH_TO_DATA>")
genotypes <- read.csv("genotypes.csv",row.names=1,col.names=TRUE)
traits <- read.csv("traits.csv",row.names=1,col.names=TRUE)
genotypes[1:5,1:10] #Show the first 5 individuals, 10 markers
traits[1:5,1:10] #Show the first 5 individuals, 10 traits
@

\emph{Use an R/qtl formatted dataset}\\
Provided is the main interface functions to R/qtl: $CTLscan.cross()$ this functions accepts and 
R/qtl formatted cross object as input and will scan CTL, perform permutations and transform the
detected differential correlation to LOD score matrices.
As an example we show the code to scan the multitrait dataset provided in R/qtl:

<<echo=TRUE,eval=FALSE>>=
require(qtl)        #Loads the R/qtl package
data(multitrait)    #Loads the dataset
multitrait          #Print basic dataset information
@

\emph{Adding genetic map information}
This step is optional, but plots and generated networks look much nicer when chromosome locations
of genetic markers are supplied. The 'mapinfo' object is matrix with 3 columns: "Chr" - the 
chromosome number, "cM" - the location of the marker in centiMorgans and the 3rd column "Mbp" - The 
location of the marker in Mega basepairs. The structure of the $mapinfo.csv$ file:

\begin{tabular}{lrrr}
      & Chr    & cM     & Mbp \\
Ind1  & 1    & 1.0   & 0.01   \\
Ind2  & 1    & 1.2   & 0.34   \\
Ind3  & 1    & 3.1   & 1.3    \\
...   & ...  & ...   & ...    \\
IndN  & 24   & 120.8 & 20.3   \\
\end{tabular}

We load the genetic map data into R by using the following command:

<<echo=TRUE,eval=FALSE>>=
mapinfo <- read.csv("mapinfo.csv",row.names=1,col.names=TRUE)
traits[1:5,1:3] #Show the first 5 marker records
@
After this we can
\bigskip

\textbf{\sffamily CTLscan and CTLscan.network}
\nopagebreak

Explains how to map CTL using the CTLscan and CTLscan.network functions.
<<echo=TRUE,eval=FALSE>>=
require(ctl)        #Loads the R/ctl package
data(ath.metab)
geno <- ath.metab$genotypes #Short name
traits <- ath.metab$phenotypes #Short name

#Scan all phenotypes for CTLs, using the default options
#NOTE: Could take some time
ctls <- CTLscan(geno,traits)
@

\emph{CTLscan options}
The genotype matrix and the $geno.enc$ param, the phenotype matrix, the important $n.perms$ param, 
which should be set to as many as possible, especially when publishing any results, but for a quick 
scan 100 is ussualy sufficient.

When experimental setups are more complex a user can supply QTL results via the $have.qtl$ option.

<<echo=TRUE,eval=FALSE>>=
require(qtl)        #load the R/qtl package
data(multitrait)    #load the dataset
qtls <- scanone(multitrait, pheno.col=1:24)[,-c(1,2)]
ctls <- CTLscan.cross(multitrait, pheno.col=1:2, have.qtl=qtls)
@

, when no qtls are supplied the internal QTL mapping routine is used, which can deal with just a 
single condition per individual. An environmental condition can be specified by the $conditions$ 
param and are also passed to the internal QTL mapping algorithm.

<<echo=TRUE,eval=FALSE>>=
#Specify that individuals 1:81 were in condition A / e.g. treatment
#Specify that individuals 82:162 were in condition B / e.g. control
conditions <- c(rep(1,81),rep(2,81))
ctls <- CTLscan(geno, traits, conditions=conditions)
@

\bigskip
\textbf{\sffamily Advanced plots and CTL networks}
\nopagebreak

So now that we have our ctls calculated we can visualize the results in multiple ways.
The easiest way is first to plot the two overview CTL heatmaps. The first heatmap shows 
the summarized LOD scores for the Trait X Marker CTLs. This plot normally looks very 
similar to a QTL heatmap:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="markers")
@
The second heatmap shows the summarized LOD scores for the Trait * Trait CTLs, this 
shows the summarized amount of evidence we find for each Trait * Trait interaction:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="phenotypes")
@
After the overview plots it is time to go more in depth, and look at some individual CTL 
profiles the CTLobject 'ctls' contains multiple CTLscans. To print a summary of the first 
scan use:
<<echo=TRUE,eval=FALSE>>=
ctls[[1]]
@
To plot other summaries just replace the 1 by the trait number. We can also use the plot 
function on individual CTLscans and shows a colorful and detailed view of the 
(selected) Trait x Marker X Traits interactions detected in the dataset.
<<echo=TRUE,eval=FALSE>>=
plot(ctls[[1]])
@

\bigskip
\textbf{\sffamily Big data and Dctl}
\nopagebreak

Explains how to use the D version to map big datsets

\bigskip
\textbf{\sffamily Significance using CTLpermute}
\nopagebreak

Explains how to determine CTL significance using the CTLpermute function.

\newpage
{\small
\noindent \textbf{\sffamily Package CTL function overview} \\
\noindent \begin{tabular}{ll} 
\hspace*{25mm} & \hspace*{103mm} \\ \hline
CTLscan & Main function for Genome wide scan for Correlated Trait Loci \\
CTLscan.network & Scan only a subset of traits for CTL using start gene(s) \\
CTLpermute & Significance thresholds for CTL by permutation\\
plot & Plot CTL profiles at increasing cut-offs \\
image & Image a multiple phenotype scan \\
\hline
\end{tabular}
}

\end{document}