### 2013-04-19 Rctl/inst/doc/manual.Rnw
Fixing issue #13, the D/test/data files are now used in the manual also

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The ctl mapping routine
%\VignetteIndexEntry{Manual}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[10pt]{article}

\usepackage{times}
\usepackage{amsmath}
\usepackage{url}
\usepackage{color}

% revise margins
\setlength{\headheight}{0.0in}
\setlength{\topmargin}{-0.25in}
\setlength{\headsep}{0.0in}
\setlength{\textheight}{9.00in}
\setlength{\footskip}{0.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}

\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}

\newcommand{\code}{\texttt}
\newcommand{\lod}{\text{LOD}}

% make "Figure" within figure captions a small font
\renewcommand{\figurename}{\small Figure}

\begin{document}

\SweaveOpts{prefix.string=fig}
\setkeys{Gin}{width=\textwidth} %% <- change width of figures
% Try to get the R code from running into the margin
<<echo=FALSE>>=
options(width=87, digits=3, scipen=4)
set.seed(61777369)
@

% Change S input/output font size
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0, 0, 0.56}}}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{fontsize=\footnotesize, baselinestretch=0.75, formatcom = {\color[rgb]{0.56, 0, 0}}}

\begin{center}
\vspace*{1in}
\textbf{\Large CTL mapping in R }

\bigskip \bigskip 
\bigskip \bigskip

{\large Danny Arends, and Ritsert C. Jansen}

\bigskip \bigskip

{\color{blue} \tt \bfseries University of Groningen \\
Groningen Bioinformatics Centre \& GCC \\
Revision \# 1 \\[24pt]
XX Okt 2011}

\end{center}

\bigskip \bigskip

\textbf{\sffamily Abstract}: Tutorial for the Correlated Trait Loci (CTL) mapping package to 
reconstruct genetic regulatory networks using the R package 'ctl' and the mapctl commandline 
tool. 
This tutorial is targetted at people with a basic understanding of genetics and would like to 
analyse there inbred cross (RIL, DH, BC) or any population using SNP markers. Main focus will 
be explaining the functionality of the software, and a how-to on using datasets currently in 
Excel, tab delimited files (.csv) or use $read.cross$ from R/qtl. In the methodology and 
background only a short introduction in CTL mapping theory is given.

\vfill
\noindent \verb|Software:      http://www.mapctl.org/download.html|\\
\noindent \verb|Manual:        http://www.mapctl.org/manual.html|\\
\noindent \verb|Source-code:   http://www.github.com/DannyArends/ctl|\\
\noindent \verb|E-mail:        Danny.Arends@Gmail.com|
\newpage

\textbf{\sffamily Some words in advance}
\nopagebreak

Welcome reader to this tutorial, you'll be thrown into the world of network reconstruction 
using differential correlations. This methodology CTL mapping is not the easiest but the 
results can be of great help unraveling trait to trait interactions in all types of 
eukaryotes. In this tutorial I'll try to use as few difficult concepts as possible. But
some always $sneak$ in.\\
\\
Enjoy this tutorial, and good luck hunting that heritability.\\
Danny Arends\\
\\
PS. Use a good text editor\\
PPS. Use a repository / backup system

\bigskip

\textbf{\sffamily Background \& Methodology}
\nopagebreak

Some introduction on the methodology, skip this chapter first (DO come back later) if you 
have a dataset and just want to use the software.

\bigskip

\textbf{\sffamily Format and load your data}
\nopagebreak

\emph{Loading your Excel / CSV data}\\
Use the example excel file and simply replace the example data by your own then: 
  $$File->Save As->TODO$$

After you have stored the data as plain text open the files in a text editor (NOT WORD, for 
windows Notepad++ is a good option) and check to see if your data is stored correctly, if 
not reformat your data in both the genotype.csv and the phenotype.csv files to match the 
format below.
{\color{red} \emph{\bf{ \\\\NOTE: Individual order must match between the genotype and the phenotype file... \\\\}}} 
The genotypes.csv file containing the genotype matrix is stored individuals x genetic marker:

\begin{tabular}{lrrrrr}
     & Chromosome & Location & Ind1 & Ind2 & Ind3 & ... & IndN \\
Marker1  & 1 & 0.0 & 1 & 1 & 1 & ... & 2 \\
marker2  & 1 & 5.3 & 1 & 2 & 2 & ... & 1 \\
marker3  & 1 & 7.8 & 2 & 2 & 1 & ... & 1 \\
...   & ... & ... & ... & ... & ... & ... & ... \\
MarkerN  & 7 & 120.1 & 1 & 1 & 1 & ... & 2 \\
\end{tabular}

The phenotypes.csv file containing individuals x trait measurements:

\begin{tabular}{lrrrrr}
      & Chromosome & Location & Ind1 & Ind2 & Ind3 & ... & IndN \\
Trait1  & ... & ... & 5.8    & 11.0   & 9.0    & ... & 1.6 \\
Trait2  & ... & ... & 6.3    & 12.2   & NA     & ... & 1.3 \\
Trait3  & ... & ... & 5.1    & 11.1   & 12.3   & ... & 2.0 \\
...   & ... & ... & ...    & ...    & ...    & ... & ... \\
Trait  & ... & ... & 9.8    & 15.8   & 23.0   & ... & 3.4 \\
\end{tabular}

After verifying your files are formatted correctly (this is the source of most errors).
Start R and load in the data, check that the output looks like below:
<<echo=TRUE,eval=FALSE>>=
  setwd("/Path/To/Data/")
  genotypes <- read.csv("genotypes.csv",row.names=1, header=FALSE,sep="\t")
  traits <- read.csv("phenotypes.csv",row.names=1, header=FALSE,sep="\t")
  genotypes[1:5,1:10]    #Show 5 individuals, 8 markers, and genetic locations
  traits[1:5,1:10]       #Show 5 individuals, 8 traits, and genetic locations
  mapinfo <- genotypes[,1:2]
  genotypes <- genotypes[,-c(1,2)]  # Remove the genetic locations
  traits <- traits[,-c(1,2)]        # Remove the genetic locations
@

\emph{Use an R/qtl formatted dataset}\\
Provided is the main interface functions to R/qtl: $CTLscan.cross()$ this functions accepts and 
R/qtl formatted cross object as input and will scan CTL, perform permutations and transform the
detected differential correlation to LOD score matrices.
As an example we show the code to load the internal multitrait dataset provided in R/qtl, load
your own by using the $read.cross()$ function:

<<echo=TRUE,eval=FALSE>>=
require(qtl)        #Loads the R/qtl package
data(multitrait)    #Loads the dataset
multitrait          #Print basic dataset information
?read.cross         #List of formats supported by R/qtl
@

\emph{Adding genetic map information}\\
This step is optional, but plots and generated networks look much nicer when chromosome locations
of genetic markers are supplied. The 'mapinfo' object is matrix with 3 columns: "Chr" - the 
chromosome number, "cM" - the location of the marker in centiMorgans and the 3rd column "Mbp" - The 
location of the marker in Mega basepairs.
{\color{red} \emph{\bf{ \\\\NOTE: Marker order must match with the genotype file... \\\\}}} 
The structure of the $mapinfo.csv$ file:

\begin{tabular}{lrrr}
      & Chr    & cM     & Mbp \\
Ind1  & 1    & 1.0   & 0.01   \\
Ind2  & 1    & 1.2   & 0.34   \\
Ind3  & 1    & 3.1   & 1.3    \\
...   & ...  & ...   & ...    \\
IndN  & 24   & 120.8 & 20.3   \\
\end{tabular}

We load the genetic map data into R by using the following command:

<<echo=TRUE,eval=FALSE>>=
mapinfo <- read.csv("mapinfo.csv",row.names=1,col.names=TRUE)
mapinfo[1:5,1:3]     #Show the first 5 marker records
@
Finally... after all this we can start scanning QTL and CTL...
\bigskip

\textbf{\sffamily Scanning for CTL / $CTLscan$}
\nopagebreak

We start explaining how to map CTL using the basic options of the $CTLscan$ function. 
This function scan CTL and produces an output "CTLobject".
<<echo=TRUE,eval=FALSE>>=
require(ctl)                     #Loads the R/ctl package
data(ath.metab)                  #Loads the example dataset
geno   <- ath.metab$genotypes    #Short name
traits <- ath.metab$phenotypes   #Short name

#Scan all phenotypes for CTLs, using the default options (Could take some time)
ctls   <- CTLscan(geno,traits)
@
If you have loaded an R/qtl "cross" object, use the $CTLscan.cross$ interface, this will 
automagically extract the cross, check if the type of "cross" is supported, and deal 
with the $geno.enc$ parameter:
<<echo=TRUE,eval=FALSE>>=
library(qtl)
data(multitrait)
ctls   <- CTLscan.cross(multitrait)
@

Now you might feel tempted to just dive into the results, But... first lets take a look at some 
options that might apply to your experimental setup like treatments applied or growth conditions, 
and which (if set incorrectly) could seriously reduce mapping power. Also you might want to add 
your own/previous QTL results.

\emph{CTLscan options}\\
The genotype matrix and the $geno.enc$ parameter, if data is of a different encoding then the 
standard 1 / 2 it can be specified by using the $geno.enc$ parameter.

<<echo=TRUE,eval=FALSE>>=
#Load a dataset with "AA" and "AB" genotypes
ctls <- CTLscan(geno,traits,geno.enc=c("AA","AB"))
@

The important $n.perms$ parameter, which should be set to as many as possible, especially 
when publishing any results, but for a quick the default of 100 is sufficient:
<<echo=TRUE,eval=FALSE>>=
ctls_quick_scan      <- CTLscan(geno,traits,n.perm=100)
ctls_for_publication <- CTLscan(geno,traits,n.perm=15000)
@

CTLs could be condition dependant just like QTL, so if one has multiple conditions or 
treatments. we need to accomodate that in the underlying models. 
{\color{red} \emph{\bf{ \\\\NOTE: Current version doesn't handle conditions (only the QTL are compensated). \\\\}}} 
If no QTLs are supplied the internal QTL mapping routine is used, which can deal with just a single condition per 
individual. An environmental condition can be specified by the $conditions$ parameter:

<<echo=TRUE,eval=FALSE>>=
#Specify that individuals 1:51 were in condition A / e.g. treatment A
#Specify that individuals 52:100 were in condition B / e.g. treatment B 
#Specify that individuals 101:162 were in condition C / e.g. control
conditions <- c(rep(1,51),rep(2,49),rep(3,62))
ctls <- CTLscan(geno, traits, conditions=conditions)
@

When experimental setups are more complex a user can supply QTL results via the $have.qtl$ option.
An example is shown below in which we use QTL results obtained by using single marker mapping via 
$scanone$ and the new Multiple QTL mapping routine $mqmscan$ in R/qtl:

<<echo=TRUE,eval=FALSE>>=
require(qtl)        #load the R/qtl package
data(multitrait)    #load the dataset
oneqtls   <- scanone(multitrait, pheno.col=1:24)[,-c(1,2)]
ctls_one  <- CTLscan.cross(multitrait, pheno.col=1:2, have.qtl=qtls)
@

And a more advanced example where the MQM scan uses 30 genetic markers as cofactors to first build 
up a genetic model underlying the QTL and locus to locus interactions:

<<echo=TRUE,eval=FALSE>>=
cofactors <- mqmautocofactors(multitrait, 30)
mqm_qtls  <- mqmscan(multitrait, cofactors=cofactors, pheno.col=1:24)[,-c(1,2)]
ctls_mqm  <- CTLscan.cross(multitrait, pheno.col=1:2, have.qtl=qtls)
@

\bigskip
\textbf{\sffamily RESULTS - Plots and CTL networks}
\nopagebreak

So now that we have our ctls calculated we can visualize the results in multiple ways.
The easiest way is first to plot the two overview CTL heatmaps. The first heatmap shows 
the summarized LOD scores for the Trait X Marker CTLs. This plot normally looks very 
similar to a QTL heatmap:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="markers")
@
The second heatmap shows the summarized LOD scores for the Trait * Trait CTLs, this 
shows the summarized amount of evidence we find for each Trait * Trait interaction:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="phenotypes")
@
After the overview plots it is time to go more in depth, and look at some individual CTL 
profiles the CTLobject 'ctls' contains multiple CTLscans. To print a summary of the first 
scan use:
<<echo=TRUE,eval=FALSE>>=
ctls[[1]]
@
To plot other summaries just replace the 1 by the trait number. We can also use the plot 
function on individual CTLscans and shows a colorful and detailed view of the 
(selected) Trait x Marker X Traits interactions detected in the dataset.
<<echo=TRUE,eval=FALSE>>=
plot(ctls[[1]])
@

\bigskip
\textbf{\sffamily Get a list of significant CTLs}
\nopagebreak

Get significant CTL interactions by using CTLsignificant.

<<echo=TRUE,eval=FALSE>>=
sign <- CTLsignificant(ctl_scan)
@

\emph{CTLnetwork and cytoscape}\\
The CTLnetwork function outputs CTL and (optional) QTL data from the CTLobject into a network 
format, if mapinfo is available it is used. The network format is relatively easy and can also
be parsed by many other programs. As an example we use cytoscape, but first lets generate some 
network input files:

<<echo=TRUE,eval=FALSE>>=
ctl_network <- CTLnetwork(ctls, lod.threshold = 5, mapinfo)
@

The CTL edges in the network\_full.sif file are annotated as:

\begin{tabular}{lrrrrr}
Trait & CTL & Trait & LOD Score\\
\end{tabular}

The CTL edges in the network\_summary.sif file look like this:

\begin{tabular}{lrrrrr}
Trait & CTL & Trait & avgLOD Score & sumLOD Score & nTraits\\
\end{tabular}

Additional QTLs are added to both files in the folowing structure:

\begin{tabular}{lrrr}
Trait & QTL & Marker & LOD Score\\
\end{tabular}

This allows the first additional infromation column (Nr 4) to act as an distance measurement, 
the closer two traits or two markers are together in the network the stronger a detected QTL 
or a CTL.

\bigskip
\textbf{\sffamily Big data and Dctl}
\nopagebreak

Big datsets require special care, R isn't the best platform to deal with large scale genomics data 
from micro- or tilling arrays (Just to mention RNAseq and full DNA strand sequencing). Thus we also
provide a commandline version of our CTL mapping routine: mapctl.exe, for more information on all 
the command line parameters can be found at http://www.mapctl.org/mapctl.html
The mapctl commandline tool uses the same input files as the R version, and supports qtab the new 
genotype and phenotype encoding scheme used by qtlHD. It provides only QTL and CTL scanning and 
permutation features, the output is analyzed in R as outlined above. To load the results from a 
mapctl scan into R by specifying the input files and the output directory created by the tool 
using:

<<echo=TRUE,eval=FALSE>>=
ctls <- read.dctl("genotypes.csv","phenotypes.csv",results="<PATH>/output/")
@

\bigskip
\textbf{\sffamily Some famous last words}
\nopagebreak

\bigskip
{\small
\noindent \textbf{\sffamily Package CTL function overview} \\
\noindent \begin{tabular}{ll} 
\hspace*{25mm} & \hspace*{103mm} \\ \hline
CTLscan & Main function for Genome wide scan for Correlated Trait Loci \\
CTLscan.network & Scan only a subset of traits for CTL using start gene(s) \\
CTLpermute & Significance thresholds for CTL by permutation\\
plot & Plot CTL profiles at increasing cut-offs \\
image & Image a multiple phenotype scan \\
\hline
\end{tabular}
}

\end{document}